---
title: "Pipeline 3: Sequences clustering"
author: "R. Espada"
output:
  html_document:
    css: style.css
    code_folding: hide
    df_print: paged
---

```{r setup, include=FALSE}
require(single)
require(dendextend)
knitr::opts_chunk$set(echo = TRUE, eval=TRUE)
knitr::opts_knit$set(root.dir = "/home/robotspm/Documents/Rocio/LRminION2020/LRminION/TestNewVersion")
```

This code reproduces the figures 3 and 4 and (supplementary figures related) in the manuscript Espada et al. 2020.

It is a pipeline to reproduce the clustering of sequences. It shows the analysis of two data sets: The first one is several barcodes, each of them with one mutant. The second one is only one barcode which includes three mutants.

Some chunks are time-consuming. So if it is the first time you run this code for a particular dataset, indicate first_time=TRUE so everything is computed and save_partial_results=TRUE to save the workspace. If not, indicate first_time=FALSE and instead of running this chunks data will be loaded. 


# General inputs
```{r general inputs, eval=TRUE}
first_time=T
save_partial_results=T


reference_sequence   = KT_wildtype
reference_sequence <- strsplit(reference_sequence, "")[[1]]

cat("Loaded reference of length: ", length(reference_sequence),"\n")
```

# 1st dataset

Inputs
```{r inputs 1, eval=TRUE}
filename_to_evaluate = c("barcode05","barcode06","barcode07","barcode08","barcode09","barcode10","barcode11")

cat("Folders to evaluate: ", filename_to_evaluate,sep="\n")
```

## Clustering

Compute matrices of distance between sequences
```{r compute distances 1}
if(first_time){
  distance_raw <- modified_hamming_distance_matrix(filename_to_evaluate, reference_sequence,weighting="raw", dist_function="dist_modified_hamming", cut_off=0.9, filename_out=NULL, filename_ids=c("05","06","07","08","09","10","11"))
  
  distance_corrected <- modified_hamming_distance_matrix(filename_to_evaluate, reference_sequence, weighting="priors", dist_function="dist_modified_hamming", cut_off=0.9, filename_out=NULL, filename_ids=c("05","06","07","08","09","10","11"))
  
  cat("done")
}else{load("Distances.Rdata");cat("Loaded")}

if(save_partial_results){ save(distance_raw, distance_corrected, file="Distances.Rdata") }
```


Compute clustering
```{r compute clustering, eval=TRUE}
k_clusters=7
if(first_time){
  raw_clustering  <- hclust(as.dist(distance_raw),  method = "ward.D2")
  raw_clusters    <- cutree(raw_clustering, k=k_clusters)
  
  corrected_clustering  <- hclust(as.dist(distance_corrected),  method = "ward.D2")
  corrected_clusters    <- cutree(corrected_clustering, k=k_clusters)
  
  corrected_ids <- parse_labels(rownames(distance_corrected), split = "_",element = 1, name=F)
  raw_ids       <- parse_labels(rownames(distance_raw), split = "_",element = 1, name=F)
 cat("done")
}else{load("Clustering.Rdata");cat("Loaded")}

if(save_partial_results){ save(raw_clustering, raw_clusters, corrected_clustering,corrected_clusters, corrected_ids,raw_ids, file="Clustering.Rdata") }
```


Plot clustering results
```{r plot clustering 1, eval=TRUE, fig.width=15}

par( mar=c(2,2,2,1), mfrow=c(1,2))
plot_tree(hierarchical_clustering = raw_clustering, ids= raw_ids, k_clusters = k_clusters, colors = c(1:7), main="Raw weights")
plot_tree(hierarchical_clustering = corrected_clustering, ids= corrected_ids, k_clusters = k_clusters, colors = c(1:7), main="Priors weights")

```



## Compute consensus by cluster

Load sequences
```{r load data for consensus 1,eval=TRUE}
if(first_time){
  data <- lapply(filename_to_evaluate, load_sequences_from_path)
  filename_ids<- c(5:11)
  ids_repetitions <- sapply(data, length)
  ids <- rep(filename_ids, ids_repetitions)
  
  if(length(filename_to_evaluate)>1){ data      <- unlist(data, recursive=FALSE) }
  
  sequences <- t(sapply(data, function(x){x$nucleotide}))
  weights_priors   <- t(sapply(data, function(x){x$p_right_priors_model})) 
  weights_raw      <- t(sapply(data, function(x){1-x$p_error_minION})) 
  cat("done")
}else{load("Data.Rdata");cat("Loaded")}

if(save_partial_results){ save(sequences,weights_priors,weights_raw, file="Data.Rdata") }
```

Calculate consensus sequence on each cluster
```{r compute consensus 1,eval=TRUE}
corrected_clusters_consensus <- clusters_consensus(sequences, corrected_clusters, reference_sequence)
corrected_clusters_consensus_weighted <- clusters_consensus(sequences, corrected_clusters, reference_sequence,weights_vector = weights_priors)

raw_clusters_consensus <- clusters_consensus(sequences, raw_clusters, reference_sequence)
raw_clusters_consensus_weighted <- clusters_consensus(sequences, raw_clusters, reference_sequence,weights_vector = weights_raw)
```

```{r,eval=TRUE}
cat("Clusters obtained with corrected weights, consensus computed without weighting")
corrected_clusters_consensus
cat("Clusters obtained with corrected weights, consensus computed weighting with p_priors")
corrected_clusters_consensus_weighted
cat("Clusters obtained with raw weights, consensus computed without weighting")
raw_clusters_consensus
cat("Clusters obtained with raw weights, consensus computed weighting with p_raw")
raw_clusters_consensus_weighted
```


# 2nd dataset

```{r inputs 2, eval=TRUE}
filename_to_evaluate_set2 = "barcode12"
```

## Clustering

Compute distance matrices
```{r compute distance 2, eval=TRUE}
if(first_time){
  distance_raw_set2 <- modified_hamming_distance_matrix(filename_to_evaluate_set2, reference_sequence,weighting="raw", 
                                   dist_function="dist_modified_hamming", cut_off=0.9,
                                   filename_out=NULL,filename_ids=NULL)
  
  distance_corrected_set2 <- modified_hamming_distance_matrix(filename_to_evaluate_set2, reference_sequence,weighting="priors", dist_function="dist_modified_hamming", cut_off=0.9,
                                    filename_out=NULL,filename_ids=NULL)
  cat("done")
}else{
  load("Distances_set2.Rdata");cat("Loaded")
}
if(save_partial_results){
  save(distance_raw_set2,distance_corrected_set2, file="Distances_set2.Rdata")
}
```


Compute clustering
```{r clustering 2, eval=TRUE}
k_clusters=3
raw_clustering_set2  <- hclust(as.dist(distance_raw_set2),  method = "ward.D2")
raw_clusters_set2    <- cutree(raw_clustering_set2, k=k_clusters)

corrected_clustering_set2  <- hclust(as.dist(distance_corrected_set2),  method = "ward.D2")
corrected_clusters_set2    <- cutree(corrected_clustering_set2, k=k_clusters)

```


Plot clustering results
```{r plot clustering 2, eval=TRUE, fig.width=15}

par( mar=c(2,2,2,1), mfrow=c(1,2))
plot_tree(hierarchical_clustering = raw_clustering_set2, ids= raw_clusters_set2, k_clusters = k_clusters, colors = c("red","blue","green"), main="Raw weights")
plot_tree(hierarchical_clustering = corrected_clustering_set2, ids= corrected_clusters_set2, k_clusters = k_clusters, colors = c("red","blue","darkgreen"), main="Priors weights")

```




## Compute consensus by cluster

Load sequences
```{r load data 2,eval=TRUE}
if(first_time){
data_set2  <- lapply(filename_to_evaluate_set2, load_sequences_from_path)

data_set2       <- unlist(data_set2 , recursive=FALSE) 
sequences_set2        <- t(sapply(data_set2 , function(x){x$nucleotide}))
weights_priors_set2   <- t(sapply(data_set2 , function(x){x$p_right_priors_model})) 
weights_raw_set2      <- t(sapply(data_set2 , function(x){1-x$p_error_minION})) 
}else{load("Data2.Rdata");cat("Loaded")}

if(save_partial_results){ save(sequences_set2,weights_priors_set2,weights_raw_set2, file="Data2.Rdata") }
```

Calculate consensus sequence on each cluster
```{r compute consensus 2,eval=TRUE}
corrected_clusters_consensus_set2 <- clusters_consensus(sequences_set2, corrected_clusters_set2, reference_sequence)
corrected_clusters_consensus_weighted_set2 <- clusters_consensus(sequences_set2, corrected_clusters_set2, reference_sequence,weights_vector = weights_priors_set2)

raw_clusters_consensus_set2 <- clusters_consensus(sequences_set2, raw_clusters_set2, reference_sequence)
raw_clusters_consensus_weighted_set2 <- clusters_consensus(sequences_set2, raw_clusters_set2, reference_sequence,weights_vector = weights_raw_set2)
```

```{r,eval=TRUE}
cat("Clusters obtained with corrected weights, consensus computed without weighting")
corrected_clusters_consensus_set2
```


```{r,eval=TRUE}
cat("Clusters obtained with corrected weights, consensus computed weighting with p_priors")
corrected_clusters_consensus_weighted_set2
```

```{r,eval=TRUE}
cat("Clusters obtained with raw weights, consensus computed without weighting")
raw_clusters_consensus_set2
```

```{r,eval=TRUE}
cat("Clusters obtained with raw weights, consensus computed weighting with p_raw")
raw_clusters_consensus_weighted_set2
```

```{r full plot clustering 2, eval=TRUE, fig.width=15}

par( mar=c(10,2,2,1), mfrow=c(1,1))
p <- plot_tree(hierarchical_clustering = corrected_clustering_set2, ids= corrected_clusters_set2, k_clusters = k_clusters, colors = c("red","blue","darkgreen"), main="Priors weights", consensus_labels = corrected_clusters_consensus_set2$consensus_mutations)

```
