---
title: "Pipeline 4: Variants' consensus"
author: "R. Espada"
output:
  html_document:
    css: style.css
    code_folding: hide
    df_print: paged
---

This code reproduces the figure 5 and (supplementary figures related) in the manuscript Espada et al. 2020.

The aim is to evaluate the improvement in the computation of the consensus sequence when using the correction presented there.

Some chunks are time-consuming. So if it is the first time you run this code for a particular dataset, indicate first_time=TRUE so everything is computed and save_partial_results=TRUE to save the workspace. If not, indicate first_time=FALSE and instead of running this chunks data will be loaded. 

```{r setup, include=FALSE, eval=TRUE}
knitr::opts_chunk$set(echo = TRUE, eval=TRUE)
knitr::opts_knit$set(root.dir = "/home/robotspm/Documents/Rocio/LRminION2020/LRminION/TestNewVersion/")
```

# Inputs
```{r inputs, warning=FALSE}
require(single)
require(bio3d);require(ggplot2)

first_time           <- TRUE
save_partial_results <- TRUE

print_auxFile        <- FALSE  #store mutations list when computing consensus in auxiliary file

reference_sequence   = KT_wildtype
reference_sequence <- strsplit(reference_sequence, "")[[1]]

# mutated variants (if know, as in this example), as a character matrix, each row a sequence
ref_seqs <- read.fasta(system.file("extdata", "Mutations_sequences.fasta", package = "minIONLogRegFull") )

# paths' names  where individual files of each sequence are stored, and the path to this paths
path_names <- c("barcode05","barcode06","barcode07","barcode08","barcode09","barcode10","barcode11")
inPath <- paste0(getwd(),"/")

# size of subsets of data to use (vector)
subsets     <- c(seq(2,10,by=2), seq(20,50,by=10))

#replicates of each subset
repetitions <- 100
```

Load racon results
```{r}

res_racoon <- read.table(system.file("extdata", "CountsRacon.txt", package = "minIONLogRegFull"))
res_racoon <- res_racoon[-1,]
res_racoon$bc <- plyr::mapvalues(res_racoon$bc,from=5:11, to=c("05","06","07","08","09","10","11"))
res_racoon <- res_racoon %>% filter(wt!="-") %>% group_by(bc, cons.nseq, seq)%>%
  dplyr::summarise(n.wrong.nucleotides=n())%>%  
  ungroup() %>%   group_by(bc,cons.nseq)%>%
  dplyr::summarise(mean_differences=as.numeric(sum(n.wrong.nucleotides)/100), 
                   sd_differences= as.numeric(sd(n.wrong.nucleotides)/100),
                   right_sequences=as.numeric(100-n()))%>%
  mutate(nseq=cons.nseq,name=paste0("barcode", bc),
         path=as.character(bc), ntest=integer(1)) %>% 
  select(-cons.nseq)%>% ungroup() %>%   select(-bc)
res_racoon$weights="Racon"
res_racoon$homopolymers="Racon"

```



# Main computation of consensus

Compute consensus in small sub-sets of sequences

```{r compute}
if(first_time){
res_priors                  <- list()
res_priors_fragment_homopol <- list()
res_priors_sort_homopol     <- list()
res_raw                  <- list()
res_raw_fragment_homopol <- list()
res_raw_sort_homopol     <- list()
barcodes<- c("05","06","07","08","09","10","11")
for(i in seq_along(barcodes) ){
  cat(i,"\n")
  res_priors[[i]] <- consensus_for_subsets(path_name = paste0(inPath,"barcode",barcodes[i],sep=""),nseq =subsets, nrep=repetitions, weighting = "priors", cutoff_prob = 0, reference_sequence = reference_sequence,reference_mutant = ref_seqs$ali[paste0("barcode",barcodes[i],sep=""),],auxFile = NULL)
  
  res_priors_fragment_homopol[[i]] <- consensus_for_subsets(path_name = paste0(inPath,"barcode",barcodes[i],sep=""),nseq =subsets, nrep=repetitions, weighting = "priors", cutoff_prob = 0, reference_sequence = reference_sequence,reference_mutant = ref_seqs$ali[paste0("barcode",barcodes[i],sep=""),],auxFile = NULL, fragment_homopolymers = T)
  
  res_priors_sort_homopol[[i]] <- consensus_for_subsets(path_name = paste0(inPath,"barcode",barcodes[i],sep=""),nseq =subsets, nrep=repetitions, weighting = "priors", cutoff_prob = 0, reference_sequence = reference_sequence,reference_mutant = ref_seqs$ali[paste0("barcode",barcodes[i],sep=""),],auxFile = NULL,sort_homopolymers = T)
  
    res_raw[[i]] <- consensus_for_subsets(path_name = paste0(inPath,"barcode",barcodes[i],sep=""),nseq =subsets, nrep=repetitions, weighting = "raw", cutoff_prob = 0, reference_sequence = reference_sequence,reference_mutant = ref_seqs$ali[paste0("barcode",barcodes[i],sep=""),],auxFile = NULL)
  
  res_raw_fragment_homopol[[i]] <- consensus_for_subsets(path_name = paste0(inPath,"barcode",barcodes[i],sep=""),nseq =subsets, nrep=repetitions, weighting = "raw", cutoff_prob = 0, reference_sequence = reference_sequence,reference_mutant = ref_seqs$ali[paste0("barcode",barcodes[i],sep=""),],auxFile = NULL, fragment_homopolymers = T)
  
  res_raw_sort_homopol[[i]] <- consensus_for_subsets(path_name = paste0(inPath,"barcode",barcodes[i],sep=""),nseq =subsets, nrep=repetitions, weighting = "raw", cutoff_prob = 0, reference_sequence = reference_sequence,reference_mutant = ref_seqs$ali[paste0("barcode",barcodes[i],sep=""),],auxFile = NULL,sort_homopolymers = T)
  
  
}

if(save_partial_results){
  save(res_priors,res_priors_fragment_homopol,res_priors_sort_homopol, 
       res_raw, res_raw_fragment_homopol, res_raw_sort_homopol, file="Consensus_auxiliary.Rdata")
  }
}else{load(file="Consensus_auxiliary.Rdata")}
```

Tidy data:

1. Label data
```{r tidy data I, warning=FALSE}
res_priors                  <- lapply(res_priors,                  function(x){x$weights="SINGLe"; x$homopolymers="Basic";return(x)})
res_priors_fragment_homopol <- lapply(res_priors_fragment_homopol, function(x){x$weights="SINGLe"; x$homopolymers="As units";return(x)})
res_priors_sort_homopol     <- lapply(res_priors_sort_homopol,     function(x){x$weights="SINGLe"; x$homopolymers="Sorted";return(x)})
res_raw                     <- lapply(res_raw,                     function(x){x$weights="Guppy"; x$homopolymers="Basic";return(x)})
res_raw_fragment_homopol    <- lapply(res_raw_fragment_homopol,    function(x){x$weights="Guppy"; x$homopolymers="As units";return(x)})
res_raw_sort_homopol        <- lapply(res_raw_sort_homopol,        function(x){x$weights="Guppy"; x$homopolymers="Sorted";return(x)})
```

2. unlist data
```{r tidy data II, warning=FALSE}
res_priors <- do.call(rbind, res_priors)
res_priors_fragment_homopol <- do.call(rbind, res_priors_fragment_homopol)
res_priors_sort_homopol <- do.call(rbind, res_priors_sort_homopol)
res_raw <- do.call(rbind, res_raw)
res_raw_fragment_homopol <- do.call(rbind, res_raw_fragment_homopol)
res_raw_sort_homopol <- do.call(rbind, res_raw_sort_homopol)
```

3. merge data
```{r tidy data III, warning=FALSE}
all_data <- rbind(res_priors,res_priors_fragment_homopol,res_priors_fragment_homopol,res_raw, res_raw_fragment_homopol, res_priors_sort_homopol)
all_data$name <- parse_labels(all_data$path, split="/", element=9, fixed=T, name = F)


all_data  = full_join(all_data,res_racoon)
```

# Plots

general aestetic values
```{r plot general}
green_fancy <- rgb(0,186,56,maxColorValue = 255)
blue_fancy  <- rgb(97,156,255,maxColorValue = 255)
red_fancy   <- rgb(248,118,109,alpha = 255, maxColorValue = 255)

v_colors          <- c(red_fancy, green_fancy, blue_fancy, grey(0.5))
names(v_colors)   <- c("Basic","Sorted","As units","Racon")
v_linetype        <- c("solid","dashed","solid")
names(v_linetype) <- c("SINGLe","Guppy","Racon")

```


## Plot how many times we get the exact correct sequence
```{r plot Nseq}
plot_all_Nseq <- ggplot(all_data,aes(x=nseq, y=right_sequences, linetype=weights, color=homopolymers))+
  geom_line()+ geom_point(size=0.7)+
  scale_color_manual(values=v_colors)+
  scale_linetype_manual(values=v_linetype)+
  facet_wrap(vars(all_data$name), nrow = 2)+
  theme_bw()+
  theme(text=element_text(size=10),legend.position = c(0.9,.22),
        legend.text = element_text(size=rel(0.5)), legend.box = "horizontal",
        legend.spacing.x = unit(0.1, "pt"),legend.spacing.y = unit(0.01,"pt"), 
        legend.key.size = unit(15,"pt"), legend.title = element_text(size=8))+
  xlab("Number of sequences averaged")+ylab("Correct consensus")+
  labs(color="Weights",linetype="Consensus")
plot_all_Nseq
```


## Plot average number of errors
```{r plot Nmut}
plot_all_Nmut <- ggplot(all_data,aes(x=nseq,  y=mean_differences, linetype=weights, color=homopolymers, ymin=mean_differences-sd_differences/2, ymax=mean_differences+sd_differences/2))+
  geom_point(size=0.7) +  geom_line() +
  scale_color_manual(values=v_colors)+
  scale_linetype_manual(values=v_linetype)+
  scale_x_continuous(breaks=seq(0,100,by=10))+
  theme_bw()+ylab("Average errors [nucleotides]")+xlab("Number of sequences averaged")+
  theme(text=element_text(size=10),legend.position = c(0.9,.15),
        legend.text = element_text(size=rel(0.5)), legend.box = "horizontal",
        legend.spacing.x = unit(0.1, "pt"),legend.spacing.y = unit(0.01,"pt"), 
        legend.key.size = unit(15,"pt"), legend.title = element_text(size=8))+
  geom_errorbar(width=0.5, linetype="solid")+
  facet_wrap(vars(all_data$name), nrow=2)+
  labs(color="Weights",linetype="Consensus")
plot_all_Nmut
```


