---
title: "Pipeline 1: Fits and modified weights"
author: "R. Espada"
output: 
  html_document:
    css: style.css
    code_folding: hide
    df_print: paged
---

This code reproduces figure 1 in the manuscript Espada et al 2020. You might need to modify lines indicated with #*

```{r setup, include=FALSE, eval=TRUE}
knitr::opts_chunk$set(echo = TRUE, eval=TRUE)
```

# Inputs
```{r packages, message=FALSE, include=FALSE, eval=TRUE}
require(single)
require(reshape2);require(ggplot2)
require(gridExtra);require(grid)
```

This are the outputs of the fittings
```{r inputs data}
data_fits <- read.table("barcode01_forward_apriori_fits.txt", header = T)  #*
data_mut <- read.table("barcode01_forward_apriori_data.txt", header = T)   #*
```
 
Important: choose position and error to plot
```{r inpute pos and base}
## Prepare data
pos <- 20               #*   Choose position to plot
mut.base <- "C"         #*   Choose nucleotide to plot
```

Choose and tidy data for selected position and error
```{r tidy data}
#Filter all data by position and base
data_chosen_pos_base <- data_mut %>% 
        filter(position==pos & nucleotide==mut.base) %>%
        mutate(ratio=counts.wt/(counts+counts.wt), ratio_scaled =counts.wt.scaled/(counts.wt.scaled+counts.scaled))
wt  <- as.character(data_chosen_pos_base$wt.base[1])

#Filter all fits' data by position and base
data_fits_chosen_pos_base <- data_fits%>%filter(position==pos & (nucleotide==mut.base | nucleotide==wt))
class_fit_prior <- sapply(data_fits_chosen_pos_base$priors, class)
if(any(class_fit_prior=="numeric")){  ind_out <- which(class_fit_prior=="numeric"); data_fits_chosen_pos_base <- data_fits_chosen_pos_base[-ind_out,] }

#Construct fitting curves
quality_range <- 0:50                         #* You'll draw the fitted curves for this Qscores (x-values)

curve_fitted_raw <- data.frame(quality=quality_range, 
                          value=glm.predict.(quality_range, slope = data_fits_chosen_pos_base$regular_slope,
                                       intercept =data_fits_chosen_pos_base$regular_intercept))

curve_fitted_corrected <- data.frame(quality=quality_range, 
                          value=glm.predict.(quality_range, slope = data_fits_chosen_pos_base$prior_slope,
                                       intercept =data_fits_chosen_pos_base$prior_intercept))

# colnames(curve_fitted_raw) <- c("quality", "value")
# curve_fitted_raw           <- melt(curve_fitted_raw, id.vars = "quality", variable.name = "nucleotide")
# 
# curve_fitted_corrected <- cbind(Qscore_range, sapply(data_fits_chosen_pos_base$priors, function(x){
#                         y <- glm.predict.(quality_range, coefficients(x)[2],coefficients(x)[1]);
#                         return(y)}))
# colnames(curve_fitted_corrected) <- c("quality", data_fits_chosen_pos_base$nucleotide)
# curve_fitted_corrected           <- melt(curve_fitted_corrected, id.vars = "quality", variable.name = "nucleotide")

```


# Plot fits
General aestetic inputs
```{r fits aestetics}
theme_fits <- theme(axis.line = element_line(linetype = 1, size = 0.),
                  text=element_text(size=10),
                  axis.text.x=element_text(size=rel(0.9)),
                  axis.text.y = element_text(size=rel(0.9)),
                  legend.text=element_text(size=rel(0.6), lineheight = 0),
                  legend.background = element_rect(fill=NA, colour = NULL),
                  legend.key.size = unit(0.2,"cm"),
                  plot.margin = unit(c(0,1,0,0), units = "mm"),
                  panel.grid = element_blank(), panel.border = element_rect(size=1))
```

plot fits:
```{r plot fits, fig.height=3, fig.width=5, message=FALSE, echo=FALSE, results='hide', warning=FALSE}
plot_fit_raw <- ggplot(data_chosen_pos_base)+
  geom_point(aes(x=quality, y=ratio), size=.5)+
  geom_line(data=curve_fitted_raw, aes(x=quality, y=value), linetype="dashed", color=red_tr)+
  scale_y_continuous(breaks=c(0,0.5,1),limits = c(0,1))+
  xlab(expression(Q[score]))+ylab(expression(n[correct]))+ggtitle("A. raw")+
  theme_bw()+theme_fits

plot_fit_corrected <- ggplot(data_chosen_pos_base)+
  geom_point(aes(x=quality, y=ratio_scaled), size=.5)+
  geom_line(data=curve_fitted_corrected, aes(x=quality, y=value), linetype="dashed", color=red_tr)+
  scale_y_continuous(breaks=c(0,0.5,1), limits = c(0,1))+
  xlab(expression(Q[score]))+ylab(expression(n[correct]))+ggtitle("B. corrected")+
  theme_bw()+theme_fits


plot_fits <- gridExtra::grid.arrange(arrangeGrob(plot_fit_raw,plot_fit_corrected, nrow=1))
plot_fits
```

# Plot histograms
General aestetic inputs
```{r hist aestetics}
red_tr <- rgb(230,0,0,alpha = 200, maxColorValue = 255)
green_tr <- rgb(0,122,0,alpha = 200, maxColorValue = 255)

theme_histograms <- theme(axis.line = element_line(linetype = 1),
                          text=element_text(size=10),
                          axis.text.x=element_text(size=rel(0.9)),
                          axis.text.y = element_text(size=rel(0.9)),
                          legend.text=element_text(size=rel(0.6), lineheight = 0),
                          legend.background = element_rect(fill=NA, colour = NULL),
                          legend.key.size = unit(0.2,"cm"),
                          plot.margin = unit(c(0,0,0,0), units = "mm"))

```

Compute histograms
```{r hist compute}
breaks <- seq(0,50,by=5)
histogram_data <- data_chosen_pos_base %>% 
  group_by(gr=cut(quality, breaks = breaks)) %>% 
  dplyr::summarise(wt= sum(counts.wt), error=sum(counts), wt.scaled=sum(counts.wt.scaled), error.scaled=sum(counts.scaled)) %>%
  mutate(x=breaks[-length(breaks)])
n <- sum(histogram_data$wt)+sum(histogram_data$error)
nscaled <- sum(histogram_data$wt.scaled)+sum(histogram_data$error.scaled)

histogram_data <- melt(histogram_data, id.vars = c("gr","x"))
histogram_data$corrected <- histogram_data$variable=="wt.scaled" | histogram_data$variable=="error.scaled" 
histogram_data$correct <- histogram_data$variable=="wt" | histogram_data$variable=="wt.scaled" 
histogram_data <- histogram_data%>%
  group_by(corrected) %>%
  dplyr::mutate(valueN=value/sum(value))
```

Plot histograms
```{r plot hist, fig.height=3, fig.width=5, message=FALSE, echo=FALSE, results='hide'}
plot_histogram_raw <- ggplot(histogram_data)+
  geom_bar(data=histogram_data%>%filter(variable=="wt"),aes(x=x,y=value/n, fill=correct), stat="identity")+
  geom_bar(data=histogram_data%>%filter(variable=="error"),aes(x=x,y=value/n, fill=correct), stat="identity", color=red_tr)+
  scale_x_continuous(breaks=round(histogram_data$x), labels = histogram_data$gr)+
  scale_fill_manual(values=c(red_tr, green_tr))+
  theme_classic()+xlab(expression(Q[score]))+ylab("Counts")+ggtitle("C Raw")+
  theme(legend.position='none')+theme_histograms+
  theme(axis.text.x=element_text(angle=90,hjust = 1, vjust = 0.5, size=rel(0.9)))

plot_histogram_corrected <- ggplot(histogram_data)+
  geom_bar(data=histogram_data%>%filter(variable=="error.scaled"),aes(x=x,y=value/nscaled, fill=correct), stat="identity")+
  geom_bar(data=histogram_data%>%filter(variable=="wt.scaled"),aes(x=x,y=value/nscaled, fill=correct), stat="identity")+
  scale_x_continuous(breaks=round(histogram_data$x), labels = histogram_data$gr)+
  scale_fill_manual(values=c(red_tr, green_tr),labels=c("Sequencing errors","Correct reads"), name="")+
  theme_classic()+xlab(expression(Q[score]))+ylab("counts")+ggtitle("D Corrected")+
  theme(legend.position = c(0.6, 1))+theme_histograms+
  theme(axis.text.x=element_text(angle=90,hjust = 1, vjust = 0.5, size=rel(0.9)))

  
plot_histograms <- gridExtra::grid.arrange(arrangeGrob(plot_histogram_raw,plot_histogram_corrected, nrow=1))
plot_histograms
#ggsave(plot_histograms, 
       #width=8, height = 4.5, units = "cm",    
       #filename = paste0("Fig_schema",pos,mut.base,"_v2.png"))
```


