---
  title: "Pipeline 3: Sequences clustering"
author: "R. Espada"
output:
  html_document:
  css: style.css
code_folding: hide
df_print: paged
---

```{r setup, include=FALSE}
require(single)
require(dendextend)
knitr::opts_chunk$set(echo = TRUE, eval=TRUE)
knitr::opts_knit$set(root.dir = "/home/robotspm/Documents/Rocio/LRminION2020/RawData/")
```

This code reproduces the figures 3 and 4 and (supplementary figures related) in the manuscript Espada et al. 2020.

It is a pipeline to reproduce the clustering of sequences. It shows the analysis of two data sets: The first one is several barcodes, each of them with one mutant. The second one is only one barcode which includes three mutants.

Some chunks are time-consuming. So if it is the first time you run this code for a particular dataset, indicate first_time=TRUE so everything is computed and save_partial_results=TRUE to save the workspace. If not, indicate first_time=FALSE and instead of running this chunks data will be loaded.


# General inputs
```{r general inputs, eval=TRUE}
first_time=F
save_partial_results=F


ref_seq   = KT_wildtype
ref_seq <- strsplit(ref_seq, "")[[1]]

cat("Loaded reference of length: ", length(ref_seq),"\n")
```

# 1st dataset

Inputs
```{r inputs 1, eval=TRUE}
filename_to_evaluate = c("barcode05","barcode06","barcode07","barcode08","barcode09","barcode10","barcode11")

cat("Files to evaluate: ", filename_to_evaluate,sep="\t")
```

## Clustering

Compute matrices of distance between sequences
```{r compute distances 1}

files_raw          <- paste0(filename_to_evaluate, "_guppy.fastq")
files_corrected    <- paste0(filename_to_evaluate, "_corrected.fastq")

if(first_time){
  ids = c("05","06","07","08","09","10","11")
  dist_fun = "dist_modified_hamming"
  #Raw
  distance_raw       <- modified_hamming_distance_matrix(filename_to_evaluate = files_raw, dist_function = dist_fun ,reference_sequence = ref_seq, cut_off = 0.9, filename_ids =ids )
  
  #Corrected
  distance_corrected <- modified_hamming_distance_matrix(filename_to_evaluate = files_corrected, dist_function = dist_fun,reference_sequence = ref_seq, cut_off = 0.9, filename_ids =ids )
  
  cat("done")
}else{load("Distances.Rdata");cat("Loaded")}

if(save_partial_results){ save(distance_raw, distance_corrected, file="Distances.Rdata") }
```


Compute clustering
```{r compute clustering, eval=TRUE}
k_clusters=7
if(first_time){
  raw_clustering  <- hclust(as.dist(distance_raw),  method = "ward.D2")
  raw_clusters    <- cutree(raw_clustering, k=k_clusters)

  corrected_clustering  <- hclust(as.dist(distance_corrected),  method = "ward.D2")
  corrected_clusters    <- cutree(corrected_clustering, k=k_clusters)

  corrected_ids <- parse_labels(rownames(distance_corrected), split = "*",element = 1, name=F, fixed=T)
  raw_ids       <- parse_labels(rownames(distance_raw), split = "*",element = 1, name=F, fixed=T)
  cat("done")
}else{load("Clustering.Rdata");cat("Loaded")}

if(save_partial_results){ save(raw_clustering, raw_clusters, corrected_clustering,corrected_clusters, corrected_ids,raw_ids, file="Clustering.Rdata") }
```


Plot clustering results
```{r plot clustering 1, eval=TRUE, fig.width=15}

par( mar=c(2,2,2,1), mfrow=c(1,2))
plot_tree(hierarchical_clustering = raw_clustering, ids= raw_ids, k_clusters = k_clusters, colors = c(1:7), main="Raw weights")
plot_tree(hierarchical_clustering = corrected_clustering, ids= corrected_ids, k_clusters = k_clusters, colors = c(1:7), main="Priors weights")

```



## Compute consensus by cluster

Load sequences
```{r load data for consensus 1,eval=TRUE}
data_corrected <- read_combine_fastq(files_corrected,identifier = c(5:11))
data_raw       <- read_combine_fastq(files_raw,identifier = c(5:11))

sequences        <- data_corrected$sequences
weights_priors   <- data_corrected$Qscores
weights_priors   <- 1 -  t(apply(mapvalues(weights_priors, from=ascii$Symbol, to=ascii$P, warn_missing = F), 1, as.numeric))
weights_raw      <- data_raw$Qscores
weights_raw      <- 1 -  t(apply(mapvalues(weights_raw, from=ascii$Symbol, to=ascii$P, warn_missing = F), 1, as.numeric))

cat("done")

```

Calculate consensus sequence on each cluster
```{r compute consensus 1,eval=TRUE}
corrected_clusters_consensus          <- clusters_consensus(sequences, corrected_clusters, ref_seq)
corrected_clusters_consensus_weighted <- clusters_consensus(sequences, corrected_clusters, ref_seq,weights_vector = weights_priors)

raw_clusters_consensus          <- clusters_consensus(sequences, raw_clusters, ref_seq)
raw_clusters_consensus_weighted <- clusters_consensus(sequences, raw_clusters, ref_seq,weights_vector = weights_raw)
```

```{r,eval=TRUE}
cat("Clusters obtained with corrected weights, consensus computed without weighting")
corrected_clusters_consensus
cat("Clusters obtained with corrected weights, consensus computed weighting with p_priors")
corrected_clusters_consensus_weighted
cat("Clusters obtained with raw weights, consensus computed without weighting")
raw_clusters_consensus
cat("Clusters obtained with raw weights, consensus computed weighting with p_raw")
raw_clusters_consensus_weighted
```
